Case study validations of automatic bullet matching

### 2018-11-13 index.Rmd
Phrasing and flow, move the aggregation section up to just after the feature/random forest discussion; keep all of the case study stuff together.

# Introduction and Background
In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with a bullet from the suspected gun under a comparison microscope. XXX this could be a second bullet from the crime scene XXX They then visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion. Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling. In other labs, CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match.

## 3D Imaging Technology
More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option. \svp{Use of a scanning electron microscope to quantitatively examine and compare bullet striae was demonstrated in \citet{ComputerIdentificationBullets1978}}. More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}. These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques. In many cases, these papers also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements. Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.

## Identification Process

@@

Quantitative features describing the similarity of two signature are extracted. Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS). We also include statistical features, such as cross-correlation and Euclidean distance between the signatures.

## Statistics for Matching Aligned Signatures
\hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter. 'roughness' defined by NIST paper. }

\svp{Cross-correlation between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where $z$ is the shift difference between the two sequences. We maximize $CC$ over values of $z$ to determine the best alignment  \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}; this maximal value is then used to assess the match strength between the two sequences. When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.}

One of the measures used by firearms examiners to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}, according to criteria used by FTEs. 
CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match. 
\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks. This can result in a consecutive matching striae of 0. } \svp{CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.}

\svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:
\begin{itemize}
\item Cross-correlation 
\item Number of matching striae 
\item Total depth of joint striations 
\item Number of non-matching striae 
\item Euclidian distance between the sequences 
\item Consecutive Matching Striae 
\item Consecutive Non-matching Striae
\end{itemize}
The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where 1 indicates that the algorithm believes it to be highly likely that the aligned signatures match, and 0 indicates that the aligned signatures are highly likely to originate from different lands.}

The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}. Set 173 was originally published as Hamby set 44. This mis-labeling has been corrected now.

The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel. While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.

# From land-to-land scores to bullet summary scores

@@

>>>>

# Introduction and Background
In current practice, firearms and toolmark examiners (FTE) evaluate the similarity of striae on bullets by placing the evidence together with another bullet under a comparison microscope. \svp{The second bullet could be a test fire from a weapon recovered during the investigation, or it could be a second bullet from the crime scene.} Examiners visually classify similarity according to the theory of firearms identification \citep{identification} as one of identification, inconclusive or exclusion. Exact guidelines for this classification vary from lab to lab; some labs will exclude only on the basis of non-matching class characteristics, such as direction of the twist in rifling, land length or number of lands, or type of rifling. In other labs, CMS (consecutively matching striae) as defined by \citet{biasotti} is used as a measure to quantify the strength of a match. \svp{In virtually all labs, individual characteristics used to identify matching bullets are derived from visual assessment; some class characteristics may be directly measured, but these are not sufficient for individualization. } <!-- Want to make it clear that individual characteristics are visually derived, rather than numerically derived... hopefully managed that without being insulting. -->

## Identification Using 3D Imaging Technology
More explicit characterization of bullet surfaces has been discussed since at least 1958 (\citeauthor{davis1968introduction}), but at the time technology was not sufficiently advanced to make an analysis based on 3D measurements or surface traces a viable option. \svp{\citet{ComputerIdentificationBullets1978} demonstrated use of a scanning electron microscope to quantitatively examine and compare bullet striae. More recently, approaches using 3D measurement data were explicitly described in 1999 (\citeauthor{dekinderAutomatedComparisonsBullet1999}), and have been further developed by \citet{bachrachDevelopment3DbasedAutomated2002,xieAutomatedBulletidentificationSystem2009, chuPilotStudyAutomated2010}. These approaches utilize 3D surface measurements directly to characterize the topology of land engraved areas (LEAs), rather than using visual or image comparison techniques. In many cases, these papers also provide some level of automation of the comparison process, with the goal of reducing human biases by augmenting the visual information with 3D measurements. Utilizing the 3D measurements allows for examination not only of the peaks in the bullet signature, but also for comparison of the valleys, which are more difficult to assess visually except in relation to the peaks.

@@

\svp{Once signatures have been extracted from stable regions with expressed striae, they must be aligned in order to assess their match strength. \citet{aoas} align signatures by maximizing the cross-correlation between the signatures. }

\svp{The cross-correlation between two signatures $S_1(x)$ and $S_2(x)$ is defined as $$CC(S_1, S_2, z) = \text{cor}\left(S_1(x), S_2(x + z)\right)$$ where $z$ is the shift difference between the two sequences. We maximize $CC$ over values of $z$ to determine the best alignment  \citep{aoas,vorburgerApplicationsCrosscorrelationFunctions2011}; this maximal value is then used to assess the match strength between the two sequences. When signatures contain missing values, pairwise complete observations are used to calculate the cross-correlation.}

## Statistics for Matching Aligned Signatures
<!-- \hh{XXX define signatures mathematically here - as random spatial process; extracted from crosscut by removing big structure first and then applying (cross-validated, \citet{aoas}) 'roughness' filter. 'roughness' defined by NIST paper. } -->

\svp{After two signatures are aligned, q}uantitative features describing the similarity of two signature are extracted. Features include characteristics of the signature that firearms and toolmark examiners base their visual assessment on, such as striae depth, number of matching striae, and the number of consecutively matching striae (CMS). Statistical features such as cross-correlation and Euclidean distance between the signatures \svp{are also extracted and used to assess the strength of the correspondance between the aligned signatures. }

One of the measures firearms examiners use to describe the strength of a match quantitatively is the number of consecutive matching striae (CMS) as defined by \citet{biasotti}.
If there are at least 6 consecutive matching striae, the bullets are considered to be a good match \citep{biasotti_firearms_1997,nichols_scientific_2006}. 
CMS in this setting are counted as the number of consecutive peaks in signatures of two lands that match\svp{, in part because it is difficult to visually assess valleys except as relative to peaks}. 
\hh{Note that because signatures are moved into phase based on the maximum cross-correlation rather than matching striation marks,}\svp{it is possible for two lands to }\hh{have a consecutive matching striae of 0. } \svp{CMS are generally considered to be a very conservative criteria, in that bullets fired through the same barrel may not have a high number of consecutively matching striae; in this sense, CMS may be more likely to produce an inconclusive result.}

\svp{The random forest presented in \citet{aoas} takes as input the following characteristics derived from a pair of aligned sequences:
\begin{itemize}
\item Cross-correlation 
\item Number of matching striae 
\item Total depth of joint striations 
\item Number of non-matching striae 
\item Euclidian distance between the sequences 
\item Consecutive Matching Striae 
\item Consecutive Non-matching Striae
\end{itemize}

The output of the random forest is a score between 0 and 1 representing the algorithm's assessment of the strength of the match, where 1 indicates that the algorithm believes it to be highly likely that the aligned signatures match, and 0 indicates that the aligned signatures are highly likely to originate from different lands.}

# Assessing Match Strength
\svp{Each of the features described in the previous section, as well as the random forest score, describe comparisons on a land-to-land basis. While these comparisons are useful, it is typically more useful to assess the entire physical object (e.g. all lands on a single bullet) to describe the correspondance between two pieces of physical evidence. In order to aggregate land-to-land match scores, we must consider the physical structure of the evidence; that is, that the sequence of lands stems from spatial information, but the numbering of each land represents the scan order and start point rather than encoding useful information. }

## Sequence Average Maximum

@@

The algorithm in \citet{aoas} is trained on scans from Hamby sets 252 and 173 made available through \citet{nist}. Set 173 was originally published as Hamby set 44. This mis-labeling has been corrected now. The scans used to train the model were taken at 20 fold magnification for a resolution of 1.5625 microns per pixel. While magnification is generally of interest in microscopy, **resolution** -- usually measured in microns per pixel -- is of more interest for 3D topographic measurements, as it determines the operative level of available data.
